---
title: "Assignment_4_262"
author: "Mikel Irigoyen & Aaron Kreisberg"
date: "2/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
# read in necessary packages (tidyverse for data manipulation/visualization and purrr for looping with pmap)
```

# Initial discussion of equation, function, and data

Here we will discuss: 
-the equation we selected. We used Manning's equation (q = (1.49/n)*(a)*(r^(2/3))*(s^(1/2)))
-the function created to perform the equation (see manning.R)
-the function's inputs, outputs, and parameters
    a = Flow Area, (ft2). Treated as constant at 3000
    n = Manning's Roughness Coefficient. Three values used (0.060, 0.065, 0.070).
    r = Hydraulic Radius, (ft) Three values used (15, 20, 25).
    s= Channel Slope, (ft/ft) 100 random values generated with a mean slope of 0.25 degrees
yielding q = Flow Rate, (ft3/s)
-data utilized for our equation
    was generated via given values and R 

# Our Data (Inputs and Parameters)

Three creeks are created with set values for Manning's roughness coefficient and channel radius

```{r sampling2, eval=TRUE, echo=TRUE}

source("../R/mannings.R")

# generate initial data structure for possible creeks

possible_creeks = data.frame(name = c("A","B","C"), n=c(0.060, 0.065, 0.070), r=c(15, 20, 25), s=.25)

# calculate a flow area 

possible_creeks$q = mannings(n = possible_creeks$n, r=possible_creeks$r, s=possible_creeks$s)

# examine data
head(possible_creeks)
```

Error checking is incorporated and tested by attempting to run the manning's equation function for a negative slope value

```{r}
possible_creeks_error = data.frame(name = c("Creek A","Creek B","Creek C"), n=c(0.040, 0.065, 0.075), r=25, s=-.25)

#Create a data frame that does not get by error checking

possible_creeks_error$q = mannings(n = possible_creeks_error$n, r=possible_creeks_error$r, s=possible_creeks_error$s)

#Error checking is set-up to return a message that slope must be greater then zero

head(possible_creeks_error)

#Error checking code in manning.r file in .R subdirectory
#mannings = function(a= 3000, r, n, s) {
#  if (s = ifelse(s < 0), return("slope must be greater then zero"), s)
#    q = (1.49/n)*(a)*(r^(2/3))*(s^(1/2))
#  return(q)
#}

```


Data is generated for slope values

```{r}
nsample = 100
mean_slope=0.25

#Generate 100 values for slope with a mean slope of 0.25 degrees

slopes = rnorm(mean=mean_slope, sd=0.1*mean_slope, nsample)
summary(slopes)

#A normal distribution is generated for slope with a standard deviation of 0.1

plot(density(slopes), ylab="Distribution of Slopes in degrees")

#Plot to assess for normal distribution

creek_flows = as.data.frame(matrix(nrow=length(slopes), ncol=nrow(possible_creeks))) 
  
# Create a data frame based on the generated slope data

head(creek_flows)


```

A for loop is constructed to generate flow rates for the slope values on three different creeks 

```{r}
nrow(possible_creeks)
for (i in 1:nrow(possible_creeks)) {
  creek_flows[,i] = mannings(n=possible_creeks$n[i], r=possible_creeks$r[i], s=slopes)
}

# Run a for loop where the 100 generated slope values are applied to the three creeks


creek_flows <- creek_flows %>% 
  mutate(slope = slopes)

#For visulization purposes ensure the slope values are included in the dataframe

head(creek_flows)

```


The resulting dataframe is plotted as using ggplot and geom_line

```{r}
for_loop_plot <- ggplot(creek_flows, aes(x=slopes)) +
  geom_line(aes(y=V1), color = "red") +
  geom_line(aes(y=V2), color = "blue") +
  geom_line(aes(y=V3), color = "orange") +
  xlab("Slope (Degrees)") +
  ylab("Flow Rate (cubic feet/second)") +
  ggtitle("For Loop Graph")
  

for_loop_plot

```


pmap from the purr package is used to perform the same looping performed by the for loop

```{r withpmap}

# the first part, generating speeds is the same
# what is I want to estimate average power use given  each car

possible_creeks

# the first part is the same as above
# use sample to generate a distribution of speeds

# assume a log normal distribution of speeds with mean 100km/hr
# recall our function needs speed in m/s not km/hr so we will also do a conversion
# lets get a sample of a 200 speeds- we could also do this by actually measuring speeds

nsample = 100
mean_slope=0.25

slopes = rnorm(mean=mean_slope, sd=0.1*mean_slope, nsample)
summary(slopes)

plot(density(slopes), ylab="Distribution of Slopes in degrees")


# how do we run each car for all speeds 
# pmap runs a function for each value in a list of parameters, with other parameters set for each iteration


creekflows_2 = pmap(list(n = possible_creeks$n, r=possible_creeks$r), mannings, s=slopes)

# lets turn to a data frame for easier graphing
# we can add column names
creekflows_2 = as.data.frame(creekflows_2, col.names=possible_creeks$name)

creekflows_2 <- creekflows_2 %>% 
  mutate(slope = slopes)

head(creekflows_2)

```



```{r}
pmap_plot <- ggplot(creekflows_2, aes(x=slope)) +
  geom_line(aes(y=A), color = "red") +
  geom_line(aes(y=B), color = "blue") +
  geom_line(aes(y=C), color = "orange") +
  xlab("Slope (Degrees)") +
  ylab("Flow Rate (cubic feet/second)") +
  ggtitle("pmap Graph")

pmap_plot
```


