---
title: "Assignment_4_262"
author: "Mikel Irigoyen & Aaron Kreisberg"
date: "2/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(purrr)
# read in necessary packages (tidyverse for data manipulation/visualization and purrr for looping with pmap)
```

# Initial discussion of equation, function, and data

Here we will discuss: 
-the equation we selected. We used Manning's equation (q = (1.49/n)*(a)*(r^(2/3))*(s^(1/2)))
-the function created to perform the equation (see manning.R)
-the function's inputs, outputs, and parameters
    a = Flow Area, (ft2)
    n = Manning's Roughness Coefficient
    r = Hydraulic Radius, (ft)
    s= Channel Slope, (ft/ft)
yielding q = Flow Rate, (ft3/s)
-data utilized for our equation
    was generated via given values and R 

# Our Data (Inputs and Parameters)



```{r sampling2, eval=TRUE, echo=TRUE}

source("../R/mannings.R")

# generate intial data structure for possible creeks

possible_creeks = data.frame(name = c("A","B","C"), n=c(0.060, 0.065, 0.070), r=c(15, 20, 25), s=.25)

# calculate a flow area 

possible_creeks$q = mannings(n = possible_creeks$n, r=possible_creeks$r, s=possible_creeks$s)

# examine data
head(possible_creeks)
```

```{r}
possible_creeks_error = data.frame(name = c("Creek A","Creek B","Creek C"), n=c(0.040, 0.065, 0.075), r=25, s=-.25)

possible_creeks_error$q = mannings(n = possible_creeks_error$n, r=possible_creeks_error$r, s=possible_creeks_error$s)

head(possible_creeks_error)

#Error checking code in manning.r file in .R subdirectory
#mannings = function(a= 3000, r, n, s) {
#  if (s = ifelse(s < 0), return("slope must be greater then zero"), s)
#    q = (1.49/n)*(a)*(r^(2/3))*(s^(1/2))
#  return(q)
#}

```



```{r}
nsample = 100
mean_slope=0.25

slopes = rnorm(mean=mean_slope, sd=0.1*mean_slope, nsample)
summary(slopes)

plot(density(slopes), ylab="Distribution of Slopes in degrees")

creek_flows = as.data.frame(matrix(nrow=length(slopes), ncol=nrow(possible_creeks)))

head(creek_flows)

nrow(possible_creeks)
for (i in 1:nrow(possible_creeks)) {
  creek_flows[,i] = mannings(n=possible_creeks$n[i], r=possible_creeks$r[i], s=slopes)
}

# now it looks like above
head(creek_flows)
```

```{r}
plot_1 = ggplot(creek_flows, aes(slopes, V1, fill=slopes))+geom_boxplot()+ggtitle("For Loop")

plot_2 = ggplot(creek_flows, aes(slopes, V2, fill=slopes))+geom_boxplot()+ggtitle("For Loop")

plot_3 = ggplot(creek_flows, aes(slopes, V3, fill=slopes))+geom_boxplot()+ggtitle("For Loop")

plot_1

plot_2

plot_3
```


```{r withpmap}

# the first part, generating speeds is the same
# what is I want to estimate average power use given  each car

possible_creeks

# the first part is the same as above
# use sample to generate a distribution of speeds

# assume a log normal distribution of speeds with mean 100km/hr
# recall our function needs speed in m/s not km/hr so we will also do a conversion
# lets get a sample of a 200 speeds- we could also do this by actually measuring speeds

nsample = 100
mean_slope=0.25

slopes = rnorm(mean=mean_slope, sd=0.1*mean_slope, nsample)
summary(slopes)

plot(density(slopes), ylab="Distribution of Slopes in degrees")


# how do we run each car for all speeds 
# pmap runs a function for each value in a list of parameters, with other parameters set for each iteration


creekflows_2 = pmap(list(n = possible_creeks$n, r=possible_creeks$r), mannings, s=slopes)

# lets turn to a data frame for easier graphing
# we can add column names
creekflows_2 = as.data.frame(creekflows_2, col.names=possible_creeks$name)

head(creekflows_2)

```

```{r}
creekflows_2 = creekflows_2 %>% pivot_longer(cols=everything(), names_to="creek", values_to="flow rate")
head(creekflows_2)


```

```{r}
method2_plot = ggplot(creekflows_2, aes(creek, flow rate, fill=creek)) + geom_boxplot()+ggtitle("pmap")

method2_plot

```


# apply family of functions does this to - FYI
# what mapply does is run the function for each row in parameters listed, using values for other parameters listed in MoreArgs EACH time - a column for row in parameter list is returned
powerall2b = mapply(FUN=autopower, A = possible_cars$area, m=possible_cars$mass, MoreArgs = list(V=speeds)  )
# we can add column names
colnames(powerall2b)=possible_cars$name

head(powerall2b)
head(powerall2)


# for plotting sometimes its useful to turn columns in to rows
powerallr2 = powerall2 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerallr2)

# quick visualization

method2_plot = ggplot(powerallr2, aes(car,power, fill=car))+geom_boxplot()+ggtitle("pmap")
method2_plot

# put plots side by side
# to confirm that they look similar
ggarrange(method1_plot, method2_plot)

# compare values
head(powerallr2)
head(powerallr1)

# not exactly the same - why?
# recall that we sample speeds!

# but we can compare means (also using  map :)
# map runs the function (here - mean), over each column 
# they are close but still not exact, so statistics such as means and quantiles are # similar, and relative difference between cars is the same
powerall1 %>% map(mean)
powerall2 %>% map(mean)

# upper 90% quantile
powerall1 %>% map(quantile, probs=0.9)
powerall2 %>% map(quantile, probs=0.9)
```
```

```{r}
# Second using R built in tools

Doing this by hand would be hard if we had many different cars - can we automate?

YES

first lets try *pmap*  

*pmap* is available in the *purr* library

*mapply* is another R option

```{r withpmap}

# the first part, generating speeds is the same
# what is I want to estimate average power use given  each car

possible_cars

# the first part is the same as above
# use sample to generate a distribution of speeds

# assume a log normal distribution of speeds with mean 100km/hr
# recall our function needs speed in m/s not km/hr so we will also do a conversion
# lets get a sample of a 200 speeds- we could also do this by actually measuring speeds

nsample = 200
mean_speed = log(100*0.277)

speeds = rlnorm(mean=mean_speed, sd=0.1*mean_speed, nsample)
summary(speeds)

plot(density(speeds), ylab="Distribution of Speeds in (m/s)")


# how do we run each car for all speeds 
# pmap runs a function for each value in a list of parameters, with other parameters set for each iteration


powerall2 = pmap(list(A = possible_cars$area, m=possible_cars$mass), autopower, V=speeds)

# lets turn to a data frame for easier graphing
# we can add column names
powerall2 = as.data.frame(powerall2, col.names=possible_cars$name)

# apply family of functions does this to - FYI
# what mapply does is run the function for each row in parameters listed, using values for other parameters listed in MoreArgs EACH time - a column for row in parameter list is returned
powerall2b = mapply(FUN=autopower, A = possible_cars$area, m=possible_cars$mass, MoreArgs = list(V=speeds)  )
# we can add column names
colnames(powerall2b)=possible_cars$name

head(powerall2b)
head(powerall2)


# for plotting sometimes its useful to turn columns in to rows
powerallr2 = powerall2 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerallr2)

# quick visualization

method2_plot = ggplot(powerallr2, aes(car,power, fill=car))+geom_boxplot()+ggtitle("pmap")
method2_plot

# put plots side by side
# to confirm that they look similar
ggarrange(method1_plot, method2_plot)

# compare values
head(powerallr2)
head(powerallr1)

# not exactly the same - why?
# recall that we sample speeds!

# but we can compare means (also using  map :)
# map runs the function (here - mean), over each column 
# they are close but still not exact, so statistics such as means and quantiles are # similar, and relative difference between cars is the same
powerall1 %>% map(mean)
powerall2 %>% map(mean)

# upper 90% quantile
powerall1 %>% map(quantile, probs=0.9)
powerall2 %>% map(quantile, probs=0.9)
```

# Third - classic looping *for*

*pmap* works quickly but it is unique to R
Other programming language (and R) use what are called loops - where repetition is more explicit

Lets do this one more time using a *for* loop

```{r withforloop}

# the first part, generating speeds is the same
# what is I want to estimate average power use given  each car

possible_creeks

# use sample to generate a distribution of speeds

# assume a log normal distribution of speeds with mean 100km/hr
# recall our function needs speed in m/s not km/hr so we will also do a conversion
# lets get a sample of a 200 speeds- we could also do this by actually measuring speeds

nsample = 200
mean_slope = log(100*0.277)

speeds = rlnorm(mean=mean_speed, sd=0.1*mean_speed, nsample)
summary(speeds)

plot(density(speeds), ylab="Distribution of Speeds in (m/s)")


# how do we run each car for all speeds 
# we use a for loop to cycle through
# we need to create a data frame to store results - as above
# one column for each car and one row for each speed

powerall3 = as.data.frame(matrix(nrow=length(speeds), ncol=nrow(possible_cars)))
# because we don't initialize it - values are NA
head(powerall3)

# how many cars area there
nrow(possible_cars)
# for loops use an index - in this case "i" but you could use anything - it repeats
# anything between the {} for each values of i between 1 and nrow(possible_car) (which is 3 in our case)

# index in to a matrix (like powerall3) is by row and column powerall3[2,5] is 2nd row and 5th column
for (i in 1:nrow(possible_cars)) {
  powerall3[,i] = autopower(A=possible_cars$area[i], m=possible_cars$mass[i], V=speeds)
}

# now it looks like above
head(powerall3)
# we can add column names
colnames(powerall3)=possible_cars$name

# plotting is the same as above

# for plotting sometimes its useful to turn columns in to rows
powerallr3 = powerall3 %>% pivot_longer(cols=everything(), names_to="car", values_to="power")
head(powerallr3)

# quick visualization
method3_plot = ggplot(powerallr3, aes(car,power, fill=car))+geom_boxplot()+ggtitle("For Loop")

ggarrange(method1_plot, method2_plot, method3_plot, nrow=3)

powerall1 %>% map(mean)
powerall2 %>% map(mean)
powerall3 %>% map(mean)

```




