---
title: "auto_power_gen_function"
author: "Aaron Kreisberg"
date: "2/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Function Example

*Input*: Vehicle speed (m/s or mps) and assuming no headwind, vehicle area (m^2), vehicle mass (kg)

*Outputs*: Automobile Power (Watts)

*Parameters*: g acceleration due to gravity (m/s^2), rho_air density of air (kg/m^3) 

What the function will do - compute power using the following equation - *body* of the function 

$P = crolling∗m∗g∗V+1/2+ρ∗cdrag∗V^3$

P is Power in watts, ρ is the density of air (~1000 $kg/m^3$), h is height in meters, r is flow rate in cubic meters per second, g is acceleration due to gravity of 9.8 $m/s^2$, $K_{Eff}$ is a coefficient of efficiency ranging from 0 to 1.

```{r}
power <- auto_power_gen_orig(speed = c(30, 22.2, 33.3, 11.1, 1.4), area = 16, mass = 31752) 

power
```

```{r}
power = as.data.frame(power)
power$speed=c(30, 22.2, 33.3, 11.1, 1.4)

ggplot(power, aes(speed, power)) + geom_line()
```

```{r}
power_3 <- auto_power_gen_orig(speed = 120*(60/1000), area = 16, mass = 31752) 

power_3
```

```{r}
power_4 <- auto_power_gen_orig(speed = 40*(60/1000), area = 16, mass = 31752) 

power_4
```

```{r}
power_6 <- auto_power_gen_orig(speed = 5*(60/1000), area = 16, mass = 31752) 

power_6
```

```{r}
powers <- (power_1, power_2, power_3, power_4, power_5, power_6)
```

